---
title: "Lab: The 12 Days of Christmas"
author: "Austin Berg"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)

# Packages for Functions
library(dplyr)
library(glue)
library(stringr)
library(purrr)
```


### A Test Dataset

Run the code below to load up a dataset called `xmas` that contains the crucial 
information about the gifts in the song. We will use this dataset to test out
our functions as we work on them.

```{r}
xmas <- read.csv("https://www.dropbox.com/s/e584pryn8evm1gz/xmas.csv?dl=1")
```

---

## Step One:  Plurals

Note that the gifts are listed in singular: for example, on day five the
narrator receives "five golden rings", but the entry in the dataset for the gift
on day 5 simply says "ring".  

Using the skeleton of the `pluralize_gift()` function, complete the code so that 
the function takes a gift and returns the appropriate plural. 

```{r}
# Function that takes a noun and makes it plural
# Arguments -- gift -- A string or vector of strings
# Return -- A string or vector of strings with the pluralized words

pluralize_gift <- function(gift){

gift_length <- str_length(gift)
  
gift <- gift %>%
  str_pad(width = gift_length+1, side = "right", pad = "s") %>% 
  str_replace_all(c("ooses" = "eese",
                    "ys" = "ies"))


return(gift)

}


```

### Test Your Function

Try your function out on the gifts in the dataset.  
(Consider: is your function *vectorized*?  It does not have to be, but it may
make things simpler!)

```{r}

pluralize_gift(xmas$Gift.Item)

purrr::map_chr(xmas$Gift.Item, pluralize_gift)

```


---


## Step Two: Creating sentences


```{r}
make_phrase <- function(num, num_word, item, verb, adjective, location) {
  
  ## Step 1: Replace NAs with blank strings
  verb <- str_replace_na(verb, "")
  adjective <- str_replace_na(adjective, "")
  location <- str_replace_na(location, "")
  
  ## Step 2: Pluralize the gift if needed
  ## Items with numbers larger than 1 should be pluralized
  ## Single items starting with a vowel should have "an" in front
  ## All other single items should have "a" in front
  
  vowel_start <- str_sub(item, 1, 1) %>% str_detect("[aeiou]")
  
  if (num > 1) {
    
    item <- pluralize_gift(item)
    
  } else if (vowel_start & num == 1) {
    item <- paste('an ', item)
    
  } else {
    item <- paste('a ', item)
    num_word <- 'and'
  }
  
  glue::glue("{num_word} {adjective} {item} {verb} {location}") %>%
    str_squish()
  
  
}
```


# Test Your Function

Try your function out on the `xmas` data, by making a new variable containing
the daily phrases.

**Hint:** The `Day.in.Words` variable isn't quite what you want! You want `12`
to say `"twelve"` **not** `"twelfth"`. I suggest using the **english** package
to create numbered days and then use those in your `pmap_chr()`!

```{r}
library(english)
xmas <- xmas %>%
  mutate(
    day_words = english::words(Day),
    day_words = ifelse(day_words == "one", "and", day_words),
    Full.Phrase = pmap_chr(list(Day, day_words, Gift.Item, 
                                Verb, Adjective, Location), 
                           make_phrase)
  ) 
```


---


## Step Three:  Iteration.

Write a function called `sing_line` that takes as input:

* A dataset

* A number indicating which day to sing about

* The name of a column in the dataset that contains the phrases for each day.

```{r}
# Produces the string for one day of the song.
# Arguments -- dataset -- A data frame containing information about gifts
# Arguments -- line -- The number of the line for the day you want to sing about
# Arguments -- phrase_col -- The variable name for the column in the dataset that contains the gift phrases
# Return -- A string singing the line of the song with all gifts for the given day.

sing_day <- function(dataset, line, phrase_col){
  
  # Step 1: Setup the intro line
  
num_word <- as.character(english::ordinal(line))
  
  intro <- glue::glue("On the {num_word} day of Christmas")
  
  # Step 2: Sing the gift phrases

    phrases <- dataset %>% 
      filter(Day <= line) %>% 
      pull({{phrase_col}}) 

    gifts <- paste(rev(phrases), collapse = '\n')
    
  ## put it together
    full_intro <- paste(intro, ', my true love sent to me,\n', sep = '')
    

  return(paste(full_intro, gifts, '\n', sep = ''))

}
```

## Step Four: Use Your Functions!

Run the following code to test out your functions! The ouput should be the 
lyrics for the entire 12 Days of Christmas song. 

```{r}
xmas2 <- xmas %>%
  mutate(
    day.num = as.character(english::english(Day)), 
    Full.Phrase = pmap_chr(list(num = Day, 
                                num_word = day.num, 
                                item = Gift.Item, 
                                verb = Verb, 
                                adjective = Adjective, 
                                location = Location), 
                           make_phrase)
  )

map_chr(1:12, ~sing_day(xmas2, .x, Full.Phrase)) %>%
  str_c(collapse = "\n") %>%
  cat()

```

# My GitHub Repository:

https://github.com/austinb787/twelvedays/tree/main


